---
title: "Clothes Price Prediction"
author: "Trevor Okinda"
date: "2024"
output: 
  html_document:
    toc: true
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---


# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        | 134780 |
| **Student Name**                             | Trevor Okinda |
| **BBIT 4.2 Group**                           | C |
| **Project Name**                             | Clothes Price Prediction |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

### Source: 

The dataset that was used can be downloaded here: *\<https://www.kaggle.com/datasets/mrsimple07/clothes-price-prediction\>*

### Reference:

*\<AAbdurakhimov, M. (2024). Clothes Price Prediction [Dataset]. Kaggle. https://www.kaggle.com/datasets/mrsimple07/clothes-price-prediction\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*

# Understanding the Dataset (Exploratory Data Analysis (EDA))
## Load dataset
```{r load dataset}
# Load necessary library
library(dplyr)

# Load dataset
clothes_data <- read.csv("clothes_price_prediction_data.csv", colClasses = c(
  Brand = "factor",
  Category = "factor",
  Color = "factor",
  Size = "factor",
  Material = "factor",
  Price = "numeric"
))

```

## Measures of Frequency
```{r MOF}
# Measures of Frequency
frequency_brand <- clothes_data %>% count(Brand)
frequency_category <- clothes_data %>% count(Category)
frequency_color <- clothes_data %>% count(Color)
frequency_size <- clothes_data %>% count(Size)
frequency_material <- clothes_data %>% count(Material)
```

## Measures of Central Tendency
```{r MOCT}
# Measures of Central Tendency
central_tendency_summary <- summarise(clothes_data, 
                                      mean_price = mean(Price),
                                      median_price = median(Price),
                                      mode_price = as.numeric(names(sort(-table(Price))[1])))

```

## Measures of Distribution
```{r MOD}
# Measures of Distribution
distribution_summary <- summarise(clothes_data,
                                  min_price = min(Price),
                                  max_price = max(Price),
                                  range_price = max_price - min_price,
                                  sd_price = sd(Price),
                                  quantiles_price = quantile(Price))
```

## Measures of Relationship
```{r MOR}
# Measures of Relationship (Correlation Matrix)
correlation_matrix <- cor(select(clothes_data, -c(1:5))) # Excluding non-numeric columns
```

## Results
```{r results}
# Display results
frequency_brand
frequency_category
frequency_color
frequency_size
frequency_material
central_tendency_summary
distribution_summary
correlation_matrix
```

## ANOVA
```{r ANOVA}
# Perform ANOVA
anova_result <- aov(Price ~ Category, data = clothes_data)

# Summary of ANOVA
summary(anova_result)
```

## Plots
```{r Plots}
# Load necessary libraries
library(ggplot2)

# Univariate Plot for Price variable
price_plot <- ggplot(clothes_data, aes(x = Price)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Distribution of Prices", x = "Price", y = "Frequency")

# Univariate Plot for Category variable
category_plot <- ggplot(clothes_data, aes(x = Category)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency of Categories", x = "Category", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display univariate plots
price_plot
category_plot
# Load necessary libraries
library(plotly)

# Multivariate Plot (Scatter Plot)
multivariate_plot <- plot_ly(data = clothes_data, x = ~Price, y = ~Category, type = "scatter", mode = "markers",
                             marker = list(color = "rgba(255, 182, 193, 0.7)", size = 10),
                             text = ~paste("Price: $", Price, "<br>Category: ", Category)) %>%
  layout(title = "Price vs. Category", xaxis = list(title = "Price"), yaxis = list(title = "Category"))

# Display multivariate plot
multivariate_plot
```

# Preprocessing and Data Transformation
## Missing Values
```{r Missing Values}
# Check for missing values
missing_values <- colSums(is.na(clothes_data))

# Print the count of missing values
print(missing_values)

```

# Training Models
## Data Splitting
```{r Data Splitting}
library(caTools)

# Split the dataset into training and testing sets
set.seed(123) # For reproducibility
split <- sample.split(clothes_data$Price, SplitRatio = 0.7)
train_data <- subset(clothes_data, split == TRUE)
test_data <- subset(clothes_data, split == FALSE)

dim(train_data)
dim(test_data)
```

## Bootstrapping
```{r Bootstrapping}
# Load necessary library
library(boot)

# Define function for bootstrapping
bootstrap_func <- function(data, indices) {
  sampled_data <- data[indices, ]
  # Perform your analysis or modeling here
  # For example, you can fit a linear model:
  lm_model <- lm(Price ~ ., data = sampled_data)
  # Then return the parameters you're interested in
  return(coef(lm_model))
}

# Perform bootstrapping
boot_results <- boot(data = clothes_data, statistic = bootstrap_func, R = 1000)

# View bootstrap results
print(boot_results)

```

## Cross-validation
```{r Cross-validation}
# Load necessary library
library(caret)

# Define cross-validation method
control <- trainControl(method = "cv", number = 10) # 10-fold cross-validation

# Train the model using cross-validation
# For example, let's say you want to train a linear regression model
model <- train(Price ~ ., data = clothes_data, method = "lm", trControl = control)

# View model results
print(model)
```

## Training Different Models
```{r Training Different Models}
# Load necessary library (if not already loaded)
library(caret)

# Define cross-validation method
control <- trainControl(method = "cv", number = 10) # 10-fold cross-validation

# Train the linear regression model using cross-validation
lm_model <- train(Price ~ ., data = clothes_data, method = "lm", trControl = control)

# View model results
print(lm_model)

# Load necessary library (if not already loaded)
library(randomForest)

# Train the random forest regression model
rf_model <- randomForest(Price ~ ., data = clothes_data)

# View model results
print(rf_model)

# Load necessary library (if not already loaded)
library(gbm)

# Train the gradient boosting regression model
gbm_model <- gbm(Price ~ ., data = clothes_data, n.trees = 100, interaction.depth = 4)

# View model results
print(gbm_model)

```

## Performance Comparison
```{r Performance Comparison}
# Load necessary libraries
library(caret)
library(randomForest)
library(gbm)

# Define cross-validation method
control <- trainControl(method = "cv", number = 10) # 10-fold cross-validation

# Define models
models <- list(
  lm = train(Price ~ ., data = clothes_data, method = "lm", trControl = control),
  rf = train(Price ~ ., data = clothes_data, method = "rf", trControl = control),
  gbm = train(Price ~ ., data = clothes_data, method = "gbm", trControl = control)
)

# Compare model performance using resamples
model_resamples <- resamples(models)

# Summarize model performance
summary(model_resamples)

# Plot model performance
dotplot(model_resamples)

```

# Saving Model
```{r Saving Model}
# Saving the Linear Regression model
saveRDS(lm_model, "./models/saved_lm_model.rds")

# Load the saved model
loaded_lm_model <- readRDS("./models/saved_lm_model.rds")

# Prepare new data for prediction
new_data <- data.frame(
  Brand = "New Balance",
  Category = "Dress",
  Color = "White",
  Size = "XS",
  Material = "Nylon"
)

# Use the loaded model to make predictions
predictions_loaded_model <- predict(loaded_lm_model, newdata = new_data)

# Print predictions
print(predictions_loaded_model)

```



